Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-16 prop_letter.tex
prop_letter.tex Add some comments...

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems~\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable byextending classical interval-valued probability
measures~\cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability
measures~\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude
that they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set
$\{\imposs,\necess\}$), respectively.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events~\cite{Kolmogorov1950}.  In the quantum case, the
events of interest are specified by \emph{projection operators}
$P$. This set $P$ includes the empty projection~$\mathbb{0}$, the
identity projection~$\mathbb{1}$, projections of the form
$\proj{\psi}$ where $\ket{\psi}$ is a pure quantum state, sums of
\emph{orthogonal} projections, and products of \emph{commuting}
projections. In a well-defined quantum probability
space~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_i$ is associated with a probability $\mu(P_i)$ subject
to the following axioms: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for a set of
mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$, we
have:
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed as
follows. First we fix a collection of intervals $\mathscr{I}$ that
represents an abstraction of the available experimental resources. At
one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x] \mid x \in [0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with ``minimal'' resources would correspond to the
following three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=[0,1]$. As before, the first
two intervals are associated with the empty projection and the
identity projection but every other projection would be associated
with $\unknown$; its probability can be anything in the range
$[0,1]$. A third special situation would be to consider IVPMs with
only the two intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$. This is what Peres calls the
\emph{crypto-deterministic} case in which there is no ambiguity about
the status of any event; we know for certain if it will occur or not.

@@

\begin{definition}
$\bar{\mu}$ is called convex~\cite{Griffiths2003}. if for all
commuting $P_{0},P_{1}\in\events$, we have
$\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$.
\end{definition}

>>>>

Our first observation is that essential objectives of our program
to achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number systems~\cite{usat,geometry2013apsrev4,DQT2014},
may be achievable by extending classical interval-valued probability
measures~\cite{JamisonLodwick2004} to the quantum domain. We begin
by axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability measures~\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on
the expectation value of the same observable expected from continuous
quantum mechanics. We then examine Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}
and the Kochen-Specker theorem in the context of QIVPM's, and conclude
that they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set~$\{\imposs,\necess\}$),
respectively.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events~\cite{Kolmogorov1950}. In the quantum case,
the events of interest are specified by \emph{projection operators}
$P$. This set $P$ includes the empty projection~$\mathbb{0}$,
the identity projection~$\mathbb{1}$, projections of the form $\proj{\psi}$
where $\ket{\psi}$ is a pure quantum state, sums of \emph{orthogonal}
projections, and products of \emph{commuting} projections. In a well-defined
quantum probability space~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_{i}$ is associated with a probability $\mu(P_{i})$
subject to the following axioms: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for a set
of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have: 
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu\left(P_{i}\right)\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\yutsung{Should we define orthogonality between projectors? Also,
in order to compare the Eq.~\ref{eq:QuantumProbability-Addition}
and convexity in definition~\ref{def:convex}, maybe it's better
to write just two terms 
\begin{equation}
\mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\textrm{ .}\label{eq:QuantumProbability-Addition-1}
\end{equation}
}

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed
as follows. First we fix a collection of intervals $\mathscr{I}$
that represents an abstraction of the available experimental resources.
At one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x]\mid x\in[0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with ``minimal'' resources would correspond to the following
three intervals $\imposs=\left[0,0\right]$, $\necess=\left[1,1\right]$,
and $\unknown=[0,1]$. As before, the first two intervals are associated
with the empty projection and the identity projection, but every other
projection would be associated with $\unknown$; its probability can
be anything in the range $[0,1]$. A third special situation would
be to consider IVPMs with only the two intervals $\imposs=\left[0,0\right]$
and $\necess=\left[1,1\right]$. This is what Peres calls the \emph{crypto-deterministic}
case~\cite{PeresRon1988,peres1995quantum} in which there is no ambiguity
about the status of any event; we know for certain if it will occur
or not.

@@

\begin{definition}\label{def:convex}$\bar{\mu}$ is called convex~\cite{Griffiths2003}
if for all commuting $P_{0},P_{1}\in\events$, we have $\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$.
\end{definition}
